Algebraic filtering of surfaces from medical images with Julia

### 2019-12-02 src/paper.tex

It is possible to show that the \textsc{lar} representation scheme is very expressive, i.e.~that it  has a large domain,  including collections of: line segments, quads, triangles, polygons, meshes;  pixels, voxels, volume images; B-reps, enumerative and decompositive representations of solids. 
In this paper we apply \textsc{lar} methods to computation of boundary representations of solid models from segmentation (labeling) of 3D medical images.
To display a triangulation of  boundary faces  in their proper position in space, the information required is contained in the \emph{geometric chain complex} (GCC):
\[
\mu: C_0\to\E^3,\ (\delta_0, \delta_1, \delta_2)
\qquad\equiv\qquad
\texttt{(geom, top) = (W, (EV, FE, CF))}
\]
The GCC allows to transform the (possibly non connected) boundary 2-cycle of surfaces as a standard B-rep~\cite{shapiroSM:202}. 
The geometry \texttt{geom} is given by
the embedding matrix \texttt{W} of vertices (0-cells), and  topology \texttt{top} by the three sparse matrices (\texttt{EV}, \texttt{FE}, \texttt{CF}) of coboundaries ($\delta_1, \delta_2, \delta_3$) of the chain complex describing a   
space arrangement~\cite{paoluzzi2019finite}.
Note that ordered pairs of letters from \texttt{V,E,F,C}, correspond to \emph{\emph{\texttt{V}}ertices$\to$\emph{\texttt{E}}dges$\to$\emph{\texttt{F}}aces$\to$\emph{\texttt{C}}ells} into the 
\emph{\emph{\texttt{C}}olumn$\to$\emph{\texttt{R}}ow} order of matrix maps of operators.

@@

\item input of 3D medical image $\mathcal{I}$ dimensions $\ell_1, \ell_2, \ell_3$, such that: $\mathcal{I} = [\ell_1]\times[\ell_2]\times[\ell_3]$, where $[\ell_k] = (1,2,\ldots,\ell_k)$;

@@

\item depending on the above, best decision for $\emph{size}$ of 3D image blocks $\mathcal{B}$ (bricks). With defaults $\emph{size}=64$,  the number of bricks will be $n=\ceil{\ell_1/size}\times\ceil{\ell_2/size}\times\ceil{\ell_3/size}$. 
Hence the default number of bricks is $n=256$, for standard images $512\times 512\times 256$;

\item computation of a Julia's sparse boundary matrix $[\partial_B]$, returning a value of type \texttt{SparseMatrixCSC\{Int8\}\{Int64\}}, where \texttt{Int8} and \texttt{Int64} are the types for values and indices, respectively, stored by Compressed Sparse Column (CSC) format; the average $[\partial_B]$ (for $\emph{size}$ = 64) is about 45 MB;

\item creation of either a local or distributed channel to implement a producer/consumer model of parallel/distributed computation, depending on available resources;

\item distribution of matrix $[\partial_B]$, of default size $45$ MB, to all available nodes/cores (workers) using the Julia macro \texttt{@eveywhere}.

@@

\subsubsection*{Job enqueuing}\label{}Communication and data synchronization may be managed through \emph{Channels}, which are the FIFO conduits that may provide producer/consumer communication. Overall execution time can be improved if other tasks can be run while a task is being executed, or while waiting for an external service/function to complete. The single work items of this stage follow:
\begin{enumerate}

@@

\item Various multiplication algorithms are being experimented, using several packages for sparse linear algebra and/or custom implementations;

@@

\item anyway, the compute elements or threads will be feed without solution of continuity in a \emph{dataflow} process. This parallel operation is, according to our preliminary experiments, the critical one of the whole workflow, in the sense that any $\Delta T$ (either positive or negative) in this stage will contribute to the total time $T$.

@@

In this paper we have introduced and discussed a Julia implementation of an algebraic filter to extract the boundary sourface of a specific image segment, described as a 3-chain of voxels, from medical 3D images. We have shown a good advantage over standard marchin-cubes algorithms. Translations from cartesian indices of cells to linearized indices, and the sparse matrix-vector multiplication are the main computational kernels of this approach. The current implementation employs Julia's channels for multiprocessing, and can be extended to gain a much greater speed-up using hybrid architectures mixing  CPUs and GPUs of last generation.

>>>>

It is possible to show that the \textsc{lar} representation scheme is very expressive, i.e.~that it  has a large domain,  including collections of: line segments, quads, triangles, polygons, meshes;  pixels, voxels, volume images; B-reps, enumerative and decompositive representations of solids. 
In this paper we apply \textsc{lar} methods to computation of boundary representations of solid models from segmentation (labeling) of 3D medical images.
To display a triangulation of  boundary faces  in their proper position in space, the information required is contained in the \emph{geometric chain complex} (GCC):
\[
\mu: C_0\to\E^3,\ (\delta_0, \delta_1, \delta_2)
\qquad\equiv\qquad
\texttt{(geom, top) = (W, (EV, FE, CF))}
\]
The GCC allows to transform the (possibly non connected) boundary 2-cycle of surfaces as a standard B-rep~\cite{shapiroSM:202}. 
The geometry \texttt{geom} is given by
the embedding matrix \texttt{W} of vertices (0-cells), and  topology \texttt{top} by the three sparse matrices (\texttt{EV}, \texttt{FE}, \texttt{CF}) of coboundaries ($\delta_0, \delta_1, \delta_2$) of the chain complex describing a   
space arrangement~\cite{paoluzzi2019finite}.
Note that ordered pairs of letters from \texttt{V,E,F,C}, correspond to \emph{\emph{\texttt{V}}ertices$\to$\emph{\texttt{E}}dges$\to$\emph{\texttt{F}}aces$\to$\emph{\texttt{C}}ells} into the 
\emph{\emph{\texttt{C}}olumn$\to$\emph{\texttt{R}}ow} order of matrix maps of operators.

@@

\item input of 3D medical image $\mathcal{I}$ with \emph{shape} $(\ell_1, \ell_2, \ell_3)$, such that: $\mathcal{I} = [\ell_1]\times[\ell_2]\times[\ell_3]$, where $[\ell_k] = [1,2,\ldots,\ell_k]$;

@@

\item depending on the above, best decision for \emph{size} of 3D image block (or brick) $\mathcal{B}$. With defaults $\emph{size}=64$,  the \emph{number of bricks} will be $n=\ceil{\ell_1/size}\times\ceil{\ell_2/size}\times\ceil{\ell_3/size}$. 
Hence the default number of bricks is $n=256$, for standard images $512\times 512\times 256$;

\item computation of the Julia's sparse boundary matrix $[\partial_B]$, returning a value of type \texttt{SparseMatrixCSC\{Int8\}\{Int64\}}, where \texttt{Int8} and \texttt{Int64} are the types for values and indices, respectively, stored by Compressed Sparse Column (CSC) format; the storage of $[\partial_B]$ (for $\emph{size}$ = 64) requires about 45 MB;

\item creation of either a local or distributed \texttt{channel} to implement a producer/consumer model of parallel/distributed computation, depending on available resources;

\item distribution of matrix $[\partial_B]$, of default size $45$ MB, to all available nodes/cores (Julia workers) using the Julia macro \texttt{@eveywhere}.

@@

\subsubsection*{Job enqueuing}\label{}
Communication and data synchronization may be managed through \emph{Channels}, which are the FIFO conduits that may provide producer/consumer communication. Overall execution time can be improved if other tasks can be run while a task is being executed, or while waiting for an external service/function to complete. The single work items of this stage follow:
\begin{enumerate}

@@

\item Various multiplication algorithms are being tested, using several packages for sparse linear algebra and/or custom implementations;

@@

\item anyway, the compute elements or threads will be fed without solution of continuity in a \emph{dataflow} process. This parallel operation is, according to our preliminary experiments, the critical one of the whole workflow, in the sense that any $\Delta T$ (either positive or negative) in this stage will contribute to the total time $T$.

@@

In this paper we have introduced and discussed a Julia implementation of an algebraic filter to extract from medical 3D images the boundary sourface of a specific image segment, described as a 3-chain of voxels. We have shown a good advantage over standard marchin-cubes algorithms. Translations from cartesian indices of cells to linearized indices, and the sparse matrix-vector multiplication are the main computational kernels of this approach. The current implementation employs Julia's channels for multiprocessing, and can be extended to gain a much greater speed-up using hybrid architectures mixing  CPUs and GPUs of last generation.
